---
title: "Reproducible Research - Assignment 1"
author: "Sergio Huerga"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

```{r getdata}
setwd("C:/Users/sghue/Documents/R")
myData<-read.csv(file="./data/activity.csv", header=TRUE)
```

## What is mean total number of steps taken per day?

```{r}
library(ggplot2)
totalSteps <- tapply(myData$steps, myData$date, FUN=sum, na.rm=TRUE)
qplot(totalSteps, binwidth=1000, xlab="Total number of steps taken each day")
meanData<-mean(totalSteps, na.rm=TRUE)
medianData<-median(totalSteps, na.rm=TRUE)
```
The mean is `r meanData` and the median is `r medianData`

## What is the average daily activity pattern?

```{r}
myAverages <- aggregate(x=list(steps=myData$steps), by=list(interval=myData$interval),
                      FUN=mean, na.rm=TRUE)
ggplot(data=myAverages, aes(x=interval, y=steps)) +
    geom_line() +
    xlab("5-minute interval") +
    ylab("average number of steps taken")
maxSteps<-myAverages[which.max(myAverages$steps),]
```

The 5-minute interval that contains the maximum number of steps is `r maxSteps`


## Imputing missing values
Note that there are a number of days/intervals where there are missing values (coded as NA). The presence of missing days may introduce bias into some calculations or summaries of the data.

```{r}
missing <- is.na(myData$steps)
table(missing)
```

All of the missing values are filled in with mean value for that 5-minute interval.

```{r}
# Replace each missing value with the mean value of its 5-minute interval
fillValue <- function(steps, interval) {
    filled <- NA
    if (!is.na(steps))
        filled <- c(steps)
    else
        filled <- (myAverages[myAverages$interval==interval, "steps"])
    return(filled)
}
filleData <- myData
filleData$steps <- mapply(fillValue, filleData$steps, filleData$interval)
```

Using the filled data set, let's make a histogram of the total number of steps taken each day and calculate the mean and median total number of steps.

```{r}
totalSteps <- tapply(filleData$steps, filleData$date, FUN=sum)
qplot(totalSteps, binwidth=1000, xlab="Total number of steps taken each day")
mean(totalSteps)
median(totalSteps)
```

Mean and median values are higher after imputing missing data. The reason is that in the original data, there are some days with `steps` values `NA` for any `interval`. 
The total number of steps taken in such days are set to 0s by default. However, after replacing missing `steps` values with the mean `steps`of associated `interval` value, these 0 values are removed from the histogram of total number of steps taken each day.

## Are there differences in activity patterns between weekdays and weekends?
According to these charts, more steps are walked dring WEEKENDS.

```{r}
weekday.weekend <- function(date) {
    day <- weekdays(date)
    if (day %in% c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday"))
        return("weekday")
    else if (day %in% c("Saturday", "Sunday"))
        return("weekend")
    else
        stop("invalid date")
}
filleData$date <- as.Date(filleData$date)
filleData$day <- sapply(filleData$date, FUN=weekday.weekend)
averages <- aggregate(steps ~ interval + day, data=filleData, mean)
ggplot(averages, aes(interval, steps)) + geom_line() + facet_grid(day ~ .) +
    xlab("5-minute interval") + ylab("Number of steps")
```

